Cognate Priming: Report

### 2021-12-06 Rmd/report.Rmd
Add Shiny app and item-level random effects to the model
Fit Bayesian model using brms

---
title: "Cognate Priming: Report"
output:
  github_document:
    pandoc_args: --webtex
always_allow_html: true
date: "`r paste0('Updated: ', format(Sys.Date(), '%d/%m/%Y'))`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = TRUE,
	warning = TRUE, 
	cache.extra = knitr::rand_seed,
	out.width = "80%",
	results = "asis",
	dpi = 25,
	dev.args = list(png = list(type = "cairo"))
)
options(
	knitr.kable.NA = '-',
	knitr.duplicate.label = "allow",
	ggplot2.discrete.fill = ggsci::pal_tron("legacy")(7),
	ggplot2.discrete.colour = ggsci::pal_tron("legacy")(7)
)

@@

# load packages
library(targets)
library(papaja)
library(knitr)
library(tidyverse)
library(lubridate)
library(scales)
library(DiagrammeR)
library(janitor)
library(gt)
library(patchwork)
library(wesanderson)
library(shiny)
library(tidybayes)
library(brms)
library(tidytext)
library(brms)
library(tidyverse)
library(here)

@@

theme_set(theme_github()) # in utils.R
```

@@

```{r participants_summary_plot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=4, out.width="100%"}
participants %>% 
	left_join(distinct(attrition, participant, valid_participant)) %>% 
	filter(valid_participant) %>% 
	drop_na(valid_participant) %>% 
	ggplot(aes(lp, fill = as.factor(test_language))) +
	facet_grid(location~age_group, scales = "free") +
	geom_bar(colour = NA) +
	geom_hline(yintercept = 24, linetype = "dashed", colour = "white") +
	labs(x = "Group", y = "Sample size", fill = "Test language") +
	theme(
		legend.position = "top",
		axis.title.x = element_blank(),
		axis.text = element_text(size = 10),
		axis.title.y = element_text(angle = 90),
		panel.grid.major.x = element_blank()
	) 
```

@@

$$

\color{white}{
\log \Bigg( \frac{p}{n - q} \Bigg) = (\beta_{0} + \beta_{0p} + \beta_{0i}) + ... \\
(\beta_{1} + \beta_{1p} + \beta_{1i}) \times Age_{pi} + ... \\
(\beta_{2} + \beta_{2p} + \beta_{2i}) \times Vocabulary_{pi} + ... \\
(\beta_{3} + \beta_{3p} + \beta_{3i})\times Condition_{pi} + ... \\
(\beta_{4} + \beta_{4i}) \times Group_{pi} + ... \\
(\beta_{5} + \beta_{5i}) \times (Vocabulary_{pi} \times Group_{pi}) + ... \\
(\beta_{6} + \beta_{6i}) \times (Vocabulary_{pi} \times Condition_{pi}) + ... \\
(\beta_{7} + \beta_{7i}) \times (Condition_{pi} \times Group_{pi}) + ... \\
(\beta_{8} + \beta_{8i}) \times (Vocabulary_{pi} \times Group_{pi} \times Condition_{pi}) + \varepsilon_{pi}
}
$$

@@

draws %>% 
	filter(!str_detect(.variable, "time_bin|sigma")) %>% 
	mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4)) %>% 
	ggplot(aes(.value, .variable, fill = stat(cut_cdf_qi(cdf, labels = percent, .width = c(0.50, 0.80, 0.95))))) +
	geom_vline(xintercept = 0, colour = "white") +
	stat_halfeye(point_color = "white", interval_color = "white") +
	labs(x = "Value", y = "Variable", fill = "CrI") +
	scale_fill_brewer(direction = -1, na.translate = FALSE) +
	theme(
		axis.title = element_blank(),
		legend.position = "top"
	)

@@

##### Overall

```{r gaze_time, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
sample_size <- participants %>% 
	filter(valid_participant) %>% 
	count(lp)

n <- expand_grid(
	lp = c("Monolingual", "Bilingual"),
	age_group = c("25 months"),
	trial_type = unique(gaze$trial_type),
	time_bin_center = seq(min(gaze$time_bin_center), max(gaze$time_bin_center), 0.5)
)

m <- epred_draws(object = fit, newdata = n, ndraws = 20, re_formula = NA) %>% 
	mutate(.value = logit_to_prob(.epred))

ggplot(m, aes(time_bin_center, .value)) +
	facet_wrap(~fct_rev(lp)) +
	geom_hline(yintercept = 0.5, colour = "white") +
	geom_line(aes(group = interaction(age_group, trial_type, lp, .draw),
				  colour = trial_type),
			  alpha = 0.15, size = 0.65) +
	stat_summary(aes(group = trial_type, colour = trial_type),
				 fun = mean, geom = "line", size = 1.25) +
	stat_summary(data = gaze, aes(y = logit_to_prob(logit_adjusted), group = trial_type,
								  colour = trial_type),
				 fun.data = mean_se, geom = "errorbar", size = 0.5, na.rm = TRUE,
				 position = position_dodge(width = 0.3)) +
	geom_text(data = sample_size, aes(x = min(gaze$time_bin_center), y = 0.95, label = paste0("N = ", n)),
			  colour = "white", hjust = 0, size = 5) +
	labs(x = "Time bin (100 ms)", y = "Expected posterior mean PTLT", colour = "Trial type",
		 fill = "Trial type", shape = "Trial type", linetype = "Trial type") +
	scale_color_brewer(palette = "Set1") +
	scale_x_continuous(breaks = min(gaze$time_bin_center):max(gaze$time_bin_center), labels = unique(gaze$time_bin)) +
	scale_y_continuous(limits = c(0, 1), labels = percent) + 
	theme(
		legend.position = "top",
		legend.title = element_blank(),
		axis.text.x = element_text(size = 9),
		panel.grid.major.x = element_blank(),
		panel.grid.minor.x = element_blank()
	)
```

>>>>

---
title: "Cognate Priming: Report"
output: html_document
always_allow_html: true
date: "`r paste0('Updated: ', format(Sys.Date(), '%d/%m/%Y'))`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = TRUE,
	warning = TRUE, 
	cache.extra = knitr::rand_seed,
	out.width = "80%",
	results = "asis",
	dpi = 300,
	dev.args = list(png = list(type = "cairo"))
)
options(
	knitr.kable.NA = '-',
	knitr.duplicate.label = "allow",
	ggplot2.discrete.fill = ggsci::pal_tron("legacy")(7),
	ggplot2.discrete.colour = ggsci::pal_tron("legacy")(7)
)

@@

@@

theme_set(theme_minimal()) # in utils.R
```

@@

```{r participants_summary_plot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=4, out.width="100%"}
participants %>% 
	left_join(distinct(attrition, participant, valid_participant)) %>% 
	filter(valid_participant) %>% 
	drop_na(valid_participant) %>% 
	ggplot(aes(lp, fill = as.factor(test_language))) +
	facet_grid(location~age_group, scales = "free") +
	geom_bar(colour = NA) +
	geom_hline(yintercept = 24, linetype = "dashed", colour = "black") +
	labs(x = "Group", y = "Sample size", fill = "Test language") +
	theme(
		legend.position = "top",
		axis.title.x = element_blank(),
		axis.text = element_text(size = 10),
		axis.title.y = element_text(angle = 90),
		panel.grid.major.x = element_blank()
	) 
```

@@

**Likelihoods**:

$$
\Bigg( \frac{p}{n - q} \Bigg) \sim \mathcal{N}(y_{ij}, ~ \sigma^2_m)
$$

**Parameters**:

$$
\begin{align}
y_{ij} =&~ (\beta_{0} + \beta_{0p} + \beta_{0i}) ~+~ ... \\ 
& (\beta_{1} + \beta_{1p} + \beta_{1i}) \times Age_{pi} ~+~ ... \\
& (\beta_{2} + \beta_{2p} + \beta_{2i}) \times Vocabulary_{pi} ~ + ~ ... \\
& (\beta_{3} + \beta_{3p} + \beta_{3i})\times Condition_{pi} ~ + ~ ... \\
& (\beta_{4} + \beta_{4i}) \times Group_{pi} ~+~ ... \\
& (\beta_{5} + \beta_{5i}) \times (Vocabulary_{pi} \times Group_{pi}) ~+~ ... \\
& (\beta_{6} + \beta_{6i}) \times (Vocabulary_{pi} \times Condition_{pi}) ~+~ ... \\
& (\beta_{7} + \beta_{7i}) \times (Condition_{pi} \times Group_{pi}) ~+~ ... \\
& (\beta_{8} + \beta_{8i}) \times (Vocabulary_{pi} \times Group_{pi} \times Condition_{pi}) ~+~ \\ & \sigma^2_{e}
\end{align}
$$

**Priors**:

$$
\begin{align}
\beta_0 \sim~ & \mathcal{N}(0, 1) \\
\beta_{1-8} \sim~ &\mathcal{N}(0, 1) \\
\sigma_m, ~\sigma^2,~ \sigma^2_p,~ \sigma^2_i \sim~ & exp(2) \\
\rho \sim ~&  LKJ(2)
\end{align}
$$

@@

draws %>% 
	filter(!str_detect(.variable, "time_bin|sigma")) %>% 
	mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4)) %>% 
	ggplot(aes(.value, .variable, fill = stat(cut_cdf_qi(cdf, labels = percent, .width = c(0.50, 0.80, 0.95))))) +
	geom_vline(xintercept = 0, colour = "black") +
	stat_halfeye(point_color = "black", interval_color = "black") +
	labs(x = "Value", y = "Variable", fill = "CrI") +
	scale_fill_brewer(direction = -1, na.translate = FALSE) +
	theme(
		axis.title = element_blank(),
		legend.position = "top"
	)

@@

##### Marginal to time course

```{r gaze_marginal, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

m <- emmeans(fit, ~trial_type:lp, by = "age_group", epred = TRUE) %>% 
	as_tibble() %>% 
	clean_names() %>% 
	mutate_at(vars(emmean:upper_hpd), function(x) ifelse(x < 0, 0, x)) %>% 
	mutate_at(vars(emmean:upper_hpd), function(x) ifelse(x > 1, 1, x))

ggplot(m, aes(age_group, emmean, ymin = lower_hpd, ymax = upper_hpd, colour = trial_type)) +
	facet_wrap(~lp, ncol = 1) +
	# geom_line(aes(group = trial_type), size = 1, position = position_dodge(width = 0.25)) +
	geom_errorbar(size = 1, width = 0, position = position_dodge(width = 0.25)) +
	geom_point(size = 4, position = position_dodge(width = 0.25)) +
	labs(x = "Age (months))", y = "P(target)", colour = "Trial type") +
	scale_y_continuous(limits = c(0, 1), labels = percent)  +
	theme(
		legend.title = element_blank(),
		legend.position = "top"
	)

##### Time course

```{r gaze_time, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
sample_size <- participants %>% 
	filter(valid_participant) %>% 
	count(lp, age_group)

n <- expand_grid(
	lp = c("Monolingual", "Bilingual"),
	age_group = paste(c(21, 25, 30), "months"),
	trial_type = unique(gaze$trial_type),
	time_bin_center = seq(min(gaze$time_bin_center), max(gaze$time_bin_center), 0.5)
)

m <- epred_draws(object = fit, newdata = n, ndraws = 50, re_formula = NA) %>% 
	mutate(.value = logit_to_prob(.epred))

ggplot(m, aes(time_bin_center, .value)) +
	facet_grid(fct_rev(lp)~age_group) +
	geom_hline(yintercept = 0.5, colour = "black") +
	geom_line(aes(group = interaction(age_group, trial_type, lp, .draw),
				  colour = trial_type),
			  alpha = 0.15, size = 0.65) +
	stat_summary(aes(group = trial_type, colour = trial_type),
				 fun = mean, geom = "line", size = 1.25) +
	stat_summary(data = gaze, aes(y = logit_to_prob(logit_adjusted), group = trial_type,
								  colour = trial_type),
				 fun.data = mean_se, geom = "errorbar", size = 0.5, na.rm = TRUE,
				 position = position_dodge(width = 0.3)) +
	geom_text(data = sample_size, aes(x = min(gaze$time_bin_center), y = 0.95, label = paste0("N = ", n)),
			  colour = "white", hjust = 0, size = 5) +
	labs(x = "Time bin (100 ms)", y = "Expected posterior mean PTLT", colour = "Trial type",
		 fill = "Trial type", shape = "Trial type", linetype = "Trial type") +
	scale_color_brewer(palette = "Set1") +
	scale_x_continuous(breaks = min(gaze$time_bin_center):max(gaze$time_bin_center), labels = unique(gaze$time_bin)) +
	scale_y_continuous(limits = c(0, 1), labels = percent) + 
	theme(
		legend.position = "top",
		legend.title = element_blank(),
		axis.text.x = element_text(size = 9),
		panel.grid.major.x = element_blank(),
		panel.grid.minor.x = element_blank()
	)
```
